---
title: "Analyze_06_PiecewiseRegression"
output: html_document
date: "2024-11-16"
---
```{r}
library(segmented)
library(microshades)
library(phyloseq)
library(vegan)
library(ggplot2)
library(dplyr)
library(tidyr)
library(metagMisc)
#library(viridis)
library(RColorBrewer)
library(ggExtra)
library(beepr)
library(readxl)
library(ggforce)
library(ggrepel)
library(ggpubr)
library(cowplot)
library(ggtext)
library(sp)
library(vegan)
library(grid)
library(ggsci)
set.seed(00110011)
theme_set(theme_bw())
theme_update(plot.title = element_text(hjust = 0.5))
depthcolors <- c("#4DAF4A", "#E41A1C", "#984EA3")
names(depthcolors) <- c("0-200m", "200-500m", "500-1000m")
markercolors <- c("#D81B60","#1E88E5")
names(markercolors) <- c("18S", "COI")

```

```{r}
load("~/Documents/Chapter4/BasinScale_Analysis_v2_2024/data_objects/CleanZooplanktonPhyloseqs.rdat")
```


# environment similarity:community similarity
- Does epipelagic/mesopelagic community similarity covary with environmental similarity? 
- Can I calculate distance among samples and test for correlation with environmental differences? 
- Is environmental similarity the same as biogeochemical provinces? Consider whether to include biogeochemical provinces

-> we'll use bray-curtis similarity, since there's variability in sampling type and depth bins etc 
##get bray curtis distances
```{r community distance}
dist_bray <- phyloseq::distance(zoops18S, method = "bray")
dist_bray_mtx18 <- as.matrix(dist_bray) # convert 'dist' to 'matrix'

# retrieve the beta-diversity value between samples:
dist_bray_mtx18["1 200-500m ", "1 500-1000m "] # 0.05152207
dist_bray_mtx18["1 200-500m ", "1 200-500m "] # 0

dist_bray_mtx_long18 <- reshape::melt(dist_bray_mtx18, as.is = TRUE)[reshape::melt(upper.tri(dist_bray_mtx18), as.is = TRUE)$value,]

dist_bray_mtx_long18$DepthA <- stringr::str_split_i((dist_bray_mtx_long18$X1), " ", i = 2)
dist_bray_mtx_long18$StationA <- stringr::str_split_i((dist_bray_mtx_long18$X1), " ", i = 1)
dist_bray_mtx_long18$DepthB <- stringr::str_split_i((dist_bray_mtx_long18$X2), " ", i = 2)
dist_bray_mtx_long18$StationB <- stringr::str_split_i((dist_bray_mtx_long18$X2), " ", i = 1)



dist_bray <- phyloseq::distance(zoopsCOI, method = "bray")
dist_bray_mtx <- as.matrix(dist_bray) # convert 'dist' to 'matrix'

# retrieve the beta-diversity value between samples:
dist_bray_mtx["1 200-500m ", "1 500-1000m "] # 0.05152207
dist_bray_mtx["1 200-500m ", "1 200-500m "] # 0

dist_bray_mtx_longCOI <- reshape::melt(dist_bray_mtx, as.is = TRUE)[reshape::melt(upper.tri(dist_bray_mtx), as.is = TRUE)$value,]

dist_bray_mtx_longCOI$DepthA <- stringr::str_split_i((dist_bray_mtx_longCOI$X1), " ", i = 2)
dist_bray_mtx_longCOI$StationA <- stringr::str_split_i((dist_bray_mtx_longCOI$X1), " ", i = 1)
dist_bray_mtx_longCOI$DepthB <- stringr::str_split_i((dist_bray_mtx_longCOI$X2), " ", i = 2)
dist_bray_mtx_longCOI$StationB <- stringr::str_split_i((dist_bray_mtx_longCOI$X2), " ", i = 1)
```

##get insitu env differences 
```{r temp sal and oxy difference}
temp <- data.frame(sample_data(zoopsCOI)[,c("MeanTemp")])
sal <- data.frame(sample_data(zoopsCOI)[,c("MeanSal")])
oxy <- data.frame(sample_data(zoopsCOI)[,c("MeanOxy")])

dist_temp <- outer(temp$MeanTemp, temp$MeanTemp, '-')
rownames(dist_temp) <- rownames(dist_bray_mtx)
colnames(dist_temp) <- colnames(dist_bray_mtx)
dist_sal <- outer(sal$MeanSal, sal$MeanSal, '-')
rownames(dist_sal) <- rownames(dist_bray_mtx)
colnames(dist_sal) <- colnames(dist_bray_mtx)
dist_oxy <- outer(oxy$MeanOxy, oxy$MeanOxy, '-')
rownames(dist_oxy) <- rownames(dist_bray_mtx)
colnames(dist_oxy) <- colnames(dist_bray_mtx)

dist_temp_coi <- reshape::melt(dist_temp, as.is = TRUE)[reshape::melt(upper.tri(dist_temp), as.is = TRUE)$value,]
dist_sal_coi <- reshape::melt(dist_sal, as.is = TRUE)[reshape::melt(upper.tri(dist_sal), as.is = TRUE)$value,]
dist_oxy_coi <- reshape::melt(dist_oxy, as.is = TRUE)[reshape::melt(upper.tri(dist_oxy), as.is = TRUE)$value,]


temp <- data.frame(sample_data(zoops18S)[,c("MeanTemp")])
sal <- data.frame(sample_data(zoops18S)[,c("MeanSal")])
oxy <- data.frame(sample_data(zoops18S)[,c("MeanOxy")])

dist_temp <- outer(temp$MeanTemp, temp$MeanTemp, '-')
rownames(dist_temp) <- rownames(dist_bray_mtx18)
colnames(dist_temp) <- colnames(dist_bray_mtx18)
dist_sal <- outer(sal$MeanSal, sal$MeanSal, '-')
rownames(dist_sal) <- rownames(dist_bray_mtx18)
colnames(dist_sal) <- colnames(dist_bray_mtx18)
dist_oxy <- outer(oxy$MeanOxy, oxy$MeanOxy, '-')
rownames(dist_oxy) <- rownames(dist_bray_mtx18)
colnames(dist_oxy) <- colnames(dist_bray_mtx18)

dist_temp_18s <- reshape::melt(dist_temp, as.is = TRUE)[reshape::melt(upper.tri(dist_temp), as.is = TRUE)$value,]
dist_sal_18s <- reshape::melt(dist_sal, as.is = TRUE)[reshape::melt(upper.tri(dist_sal), as.is = TRUE)$value,]
dist_oxy_18s <- reshape::melt(dist_oxy, as.is = TRUE)[reshape::melt(upper.tri(dist_oxy), as.is = TRUE)$value,]
```

##get model env differences 
```{r temp sal and oxy difference}
temp <- data.frame(sample_data(zoopsCOI)[,c("meanModelTemp")])
sal <- data.frame(sample_data(zoopsCOI)[,c("meanModelSal")])

dist_temp <- outer(temp$meanModelTemp, temp$meanModelTemp, '-')
rownames(dist_temp) <- rownames(dist_bray_mtx)
colnames(dist_temp) <- colnames(dist_bray_mtx)
dist_sal <- outer(sal$meanModelSal, sal$meanModelSal, '-')
rownames(dist_sal) <- rownames(dist_bray_mtx)
colnames(dist_sal) <- colnames(dist_bray_mtx)

dist_modtemp_coi <- reshape::melt(dist_temp, as.is = TRUE)[reshape::melt(upper.tri(dist_temp), as.is = TRUE)$value,]
dist_modsal_coi <- reshape::melt(dist_sal, as.is = TRUE)[reshape::melt(upper.tri(dist_sal), as.is = TRUE)$value,]


temp <- data.frame(sample_data(zoops18S)[,c("meanModelTemp")])
sal <- data.frame(sample_data(zoops18S)[,c("meanModelSal")])

dist_temp <- outer(temp$meanModelTemp, temp$meanModelTemp, '-')
rownames(dist_temp) <- rownames(dist_bray_mtx18)
colnames(dist_temp) <- colnames(dist_bray_mtx18)
dist_sal <- outer(sal$meanModelSal, sal$meanModelSal, '-')
rownames(dist_sal) <- rownames(dist_bray_mtx18)
colnames(dist_sal) <- colnames(dist_bray_mtx18)

dist_modtemp_18s <- reshape::melt(dist_temp, as.is = TRUE)[reshape::melt(upper.tri(dist_temp), as.is = TRUE)$value,]
dist_modsal_18s <- reshape::melt(dist_sal, as.is = TRUE)[reshape::melt(upper.tri(dist_sal), as.is = TRUE)$value,]
```
 

##get food env differences 
```{r temp sal and oxy difference}
npp <- data.frame(sample_data(zoopsCOI)[,c("npp_meanyearlytotal")])
ef <- data.frame(sample_data(zoopsCOI)[,c("ef_meanyearlytotal")])

dist_npp <- outer(npp$npp_meanyearlytotal, npp$npp_meanyearlytotal, '-')
rownames(dist_npp) <- rownames(dist_bray_mtx)
colnames(dist_npp) <- colnames(dist_bray_mtx)
dist_ef <- outer(ef$ef_meanyearlytotal, ef$ef_meanyearlytotal, '-')
rownames(dist_ef) <- rownames(dist_bray_mtx)
colnames(dist_ef) <- colnames(dist_bray_mtx)

dist_npp_coi <- reshape::melt(dist_npp, as.is = TRUE)[reshape::melt(upper.tri(dist_npp), as.is = TRUE)$value,]
dist_ef_coi <- reshape::melt(dist_ef, as.is = TRUE)[reshape::melt(upper.tri(dist_ef), as.is = TRUE)$value,]


npp <- data.frame(sample_data(zoops18S)[,c("npp_meanyearlytotal")])
ef <- data.frame(sample_data(zoops18S)[,c("ef_meanyearlytotal")])

dist_temp <- outer(npp$npp_meanyearlytotal, npp$npp_meanyearlytotal, '-')
rownames(dist_temp) <- rownames(dist_bray_mtx18)
colnames(dist_temp) <- colnames(dist_bray_mtx18)
dist_ef <- outer(ef$ef_meanyearlytotal, ef$ef_meanyearlytotal, '-')
rownames(dist_ef) <- rownames(dist_bray_mtx18)
colnames(dist_ef) <- colnames(dist_bray_mtx18)

dist_npp_18s <- reshape::melt(dist_npp, as.is = TRUE)[reshape::melt(upper.tri(dist_npp), as.is = TRUE)$value,]
dist_ef_18s <- reshape::melt(dist_ef, as.is = TRUE)[reshape::melt(upper.tri(dist_ef), as.is = TRUE)$value,]
```
 
 

##reformat 18S for regression
```{r}
dist_modtemp_18s$variable <- "Temperature"
dist_modsal_18s$variable <- "Salinity"
dist_oxy_18s$variable <- "Oxygen"
dist_ef_18s$variable <- "EF"
dist_npp_18s$variable <- "NPP"
colnames(dist_modtemp_18s)[3] <- "Temperature"
colnames(dist_modsal_18s)[3] <- "Salinity"
colnames(dist_oxy_18s)[3] <- "Oxygen"
colnames(dist_ef_18s)[3] <- "EF"
colnames(dist_npp_18s)[3] <- "NPP"
dist_ef_18s$EF <- dist_ef_18s$EF * 365
dist_npp_18s$NPP <- dist_npp_18s$NPP * 365
dist_modtemp_18s$formerging <- paste(dist_modtemp_18s$X1, dist_modtemp_18s$X2)
dist_modsal_18s$formerging <- paste(dist_modsal_18s$X1, dist_modsal_18s$X2)
dist_oxy_18s$formerging <- paste(dist_oxy_18s$X1, dist_oxy_18s$X2)
dist_ef_18s$formerging <- paste(dist_ef_18s$X1, dist_ef_18s$X2)
dist_npp_18s$formerging <- paste(dist_npp_18s$X1, dist_npp_18s$X2)


colnames(dist_bray_mtx_long18)[3] <- "BCD"
dist_bray_mtx_long18$formerging <- paste(dist_bray_mtx_long18$X1, dist_bray_mtx_long18$X2)
dist_bray_mtx_long18 <- dist_bray_mtx_long18[dist_bray_mtx_long18$DepthA == dist_bray_mtx_long18$DepthB,]
dist_bray_mtx_long18 <- merge(dist_bray_mtx_long18, dist_modtemp_18s, by = "formerging", all.x = T, all.y = F) %>% 
  merge(dist_modsal_18s, by = "formerging", all.x = T, all.y = F)%>% 
  merge(dist_oxy_18s, by = "formerging", all.x = T, all.y = F) %>% 
  merge(dist_ef_18s, by = "formerging", all.x = T, all.y = F) %>% 
  merge(dist_npp_18s, by = "formerging", all.x = T, all.y = F)
forregression_18smod <- dist_bray_mtx_long18 %>% dplyr::select(DepthA, StationA, DepthB, StationB, BCD, Temperature, Salinity, Oxygen, EF, NPP) %>% pivot_longer(-c(DepthA, StationA, DepthB, StationB, BCD), values_to = "EnvDiff")
forregression_18smod$EnvDiff <- abs(forregression_18smod$EnvDiff)
forregression_18smod$name <- forcats::fct_relevel(forregression_18smod$name, "Temperature", "Salinity", "Oxygen", "NPP", "EF")
forregression_18smod$name_fac <- factor(forregression_18smod$name, levels = c("Temperature", "Salinity", "Oxygen", "NPP", "EF"), ordered = TRUE, labels = c(expression(Temperature~(degree*C)), "Salinity", expression(DO~(µmol~kg^-1)), expression(NPP~(mgC~m^-2~yr^1)), expression(EF~(mgC~m^-2~yr^1))))

```



##reformat COI for regression
```{r}

dist_modtemp_coi$variable <- "Temperature"
dist_modsal_coi$variable <- "Salinity"
dist_oxy_coi$variable <- "Oxygen"
dist_ef_coi$variable <- "EF"
dist_npp_coi$variable <- "NPP"
colnames(dist_modtemp_coi)[3] <- "Temperature"
colnames(dist_modsal_coi)[3] <- "Salinity"
colnames(dist_oxy_coi)[3] <- "Oxygen"
colnames(dist_ef_coi)[3] <- "EF"
colnames(dist_npp_coi)[3] <- "NPP"
dist_ef_coi$EF <- dist_ef_coi$EF * 365
dist_npp_coi$NPP <- dist_npp_coi$NPP * 365
dist_modtemp_coi$formerging <- paste(dist_modtemp_coi$X1, dist_modtemp_coi$X2)
dist_modsal_coi$formerging <- paste(dist_modsal_coi$X1, dist_modsal_coi$X2)
dist_oxy_coi$formerging <- paste(dist_oxy_coi$X1, dist_oxy_coi$X2)
dist_ef_coi$formerging <- paste(dist_ef_coi$X1, dist_ef_coi$X2)
dist_npp_coi$formerging <- paste(dist_npp_coi$X1, dist_npp_coi$X2)

colnames(dist_bray_mtx_longCOI)[3] <- "BCD"
dist_bray_mtx_longCOI$formerging <- paste(dist_bray_mtx_longCOI$X1, dist_bray_mtx_longCOI$X2)
dist_bray_mtx_longCOI <- dist_bray_mtx_longCOI[dist_bray_mtx_longCOI$DepthA == dist_bray_mtx_longCOI$DepthB,]
dist_bray_mtx_longCOI <- merge(dist_bray_mtx_longCOI, dist_modtemp_coi, by = "formerging", all.x = T, all.y = F) %>% 
  merge(dist_modsal_coi, by = "formerging", all.x = T, all.y = F)%>% 
  merge(dist_oxy_coi, by = "formerging", all.x = T, all.y = F)%>% 
  merge(dist_ef_coi, by = "formerging", all.x = T, all.y = F)%>% 
  merge(dist_npp_coi, by = "formerging", all.x = T, all.y = F)
forregression_COImod <- dist_bray_mtx_longCOI %>% dplyr::select(DepthA, StationA, DepthB, StationB, BCD, Temperature, Salinity, Oxygen, EF, NPP) %>% pivot_longer(-c(DepthA, StationA, DepthB, StationB, BCD), values_to = "EnvDiff")
forregression_COImod$EnvDiff <- abs(forregression_COImod$EnvDiff)
forregression_COImod$name <- forcats::fct_relevel(forregression_COImod$name, "Temperature", "Salinity", "Oxygen", "NPP", "EF")
forregression_COImod$name_fac <- factor(forregression_COImod$name, levels = c("Temperature", "Salinity", "Oxygen", "NPP", "EF"), ordered = TRUE, labels = c(expression(Temperature~(degree*C)), "Salinity", expression(DO~(µmol~kg^-1)), expression(NPP~(mgC~m^-2~yr^1)), expression(EF~(mgC~m^-2~yr^1))))
```


##piecewise 18S
```{r}
forregression_18smodwide <- forregression_18smod %>% pivot_wider(
  names_from = name, 
  values_from = EnvDiff, 
  id_cols = contains(c("Depth", "Station", "BCD")), 
  values_fn = ~ mean(.x, na.rm = TRUE))
#fit simple linear regression model
fit <- lm(BCD ~ Temperature, data=forregression_18smodwide) #Salinity + NPP + Oxygen

#fit piecewise regression model to original model, estimating a breakpoint at x=9
segmented.fit <- segmented(fit, seg.Z = ~Temperature, psi = c(Temperature = 6))
# + Salinity + NPP + Oxygen
#, Salinity = 3, NPP = 3, Oxygen = 3

#view summary of segmented model
summary(segmented.fit)

plot(forregression_18smodwide$Temperature, forregression_18smodwide$BCD, pch=16, col='steelblue')

#add segmented regression model
plot(segmented.fit, add=T)
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```

