---
title: "Import_05_CalculateClimatologyEnvironment"
output: html_document
date: "2024-10-31"
---

#load libraries
```{r}
library(ncdf4) # package for netcdf manipulation
library(raster) # package for raster manipulation
library(rgdal) # package for geospatial analysis
library(ggplot2) # package for plotting
```

Note: salinity: salinity[psu] = sali*1000+35
Note: depth: 2.50000000000 to 5900.00000000000 (54 points, avg. res. 111.274)

#SALINITY
##get data
```{r}
nc_data <- nc_open('/Users/stephaniematthews/ofesdata/sali00000.nc')
lon <- ncvar_get(nc_data, "lon")
lat <- ncvar_get(nc_data, "lat", verbose = F)
salinity1 <- ncvar_get(nc_data, "sali") # store the data in a 3-dimensional array
fillvalue <- ncatt_get(nc_data, "sali", "_FillValue")
levels_m <- ncvar_get(nc_data, "lev")
fillvalue
nc_close(nc_data) 
nc_data <- nc_open('/Users/stephaniematthews/ofesdata/sali00001.nc')
salinity2 <- ncvar_get(nc_data, "sali") # store the data in a 3-dimensional array
nc_close(nc_data) 
nc_data <- nc_open('/Users/stephaniematthews/ofesdata/sali00002.nc')
salinity3 <- ncvar_get(nc_data, "sali") # store the data in a 3-dimensional array
nc_close(nc_data) 
nc_data <- nc_open('/Users/stephaniematthews/ofesdata/sali00003.nc')
salinity4 <- ncvar_get(nc_data, "sali") # store the data in a 3-dimensional array
nc_close(nc_data) 
nc_data <- nc_open('/Users/stephaniematthews/ofesdata/sali00004.nc')
salinity5 <- ncvar_get(nc_data, "sali") # store the data in a 3-dimensional array
nc_close(nc_data) 
nc_data <- nc_open('/Users/stephaniematthews/ofesdata/sali00005.nc')
salinity6 <- ncvar_get(nc_data, "sali") # store the data in a 3-dimensional array
nc_close(nc_data) 
nc_data <- nc_open('/Users/stephaniematthews/ofesdata/sali00006.nc')
salinity7 <- ncvar_get(nc_data, "sali") # store the data in a 3-dimensional array
nc_close(nc_data) 
nc_data <- nc_open('/Users/stephaniematthews/ofesdata/sali00007.nc')
salinity8 <- ncvar_get(nc_data, "sali") # store the data in a 3-dimensional array
nc_close(nc_data) 
nc_data <- nc_open('/Users/stephaniematthews/ofesdata/sali00008.nc')
salinity9 <- ncvar_get(nc_data, "sali") # store the data in a 3-dimensional array
nc_close(nc_data) 
nc_data <- nc_open('/Users/stephaniematthews/ofesdata/sali00009.nc')
salinity10 <- ncvar_get(nc_data, "sali") # store the data in a 3-dimensional array
nc_close(nc_data) 
nc_data <- nc_open('/Users/stephaniematthews/ofesdata/sali00010.nc')
salinity11 <- ncvar_get(nc_data, "sali") # store the data in a 3-dimensional array
nc_close(nc_data) 
nc_data <- nc_open('/Users/stephaniematthews/ofesdata/sali00011.nc')
salinity12 <- ncvar_get(nc_data, "sali") # store the data in a 3-dimensional array
nc_close(nc_data) 
```

##reassign NA value
```{r}
salinity1[salinity1 == fillvalue$value] <- NA
salinity2[salinity2 == fillvalue$value] <- NA
salinity3[salinity3 == fillvalue$value] <- NA
salinity4[salinity4 == fillvalue$value] <- NA
salinity5[salinity5 == fillvalue$value] <- NA
salinity6[salinity6 == fillvalue$value] <- NA
salinity7[salinity7 == fillvalue$value] <- NA
salinity8[salinity8 == fillvalue$value] <- NA
salinity9[salinity9 == fillvalue$value] <- NA
salinity10[salinity10 == fillvalue$value] <- NA
salinity11[salinity11 == fillvalue$value] <- NA
salinity12[salinity12 == fillvalue$value] <- NA
```

##format as brick
```{r}
r_brick1 <- brick(salinity1, xmn=min(lat), xmx=max(lat), ymn=min(lon), ymx=max(lon), crs=CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0"))
r_brick2 <- brick(salinity2, xmn=min(lat), xmx=max(lat), ymn=min(lon), ymx=max(lon), crs=CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0"))
r_brick3 <- brick(salinity3, xmn=min(lat), xmx=max(lat), ymn=min(lon), ymx=max(lon), crs=CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0"))
r_brick4 <- brick(salinity4, xmn=min(lat), xmx=max(lat), ymn=min(lon), ymx=max(lon), crs=CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0"))
r_brick5 <- brick(salinity5, xmn=min(lat), xmx=max(lat), ymn=min(lon), ymx=max(lon), crs=CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0"))
r_brick6 <- brick(salinity6, xmn=min(lat), xmx=max(lat), ymn=min(lon), ymx=max(lon), crs=CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0"))
r_brick7 <- brick(salinity7, xmn=min(lat), xmx=max(lat), ymn=min(lon), ymx=max(lon), crs=CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0"))
r_brick8 <- brick(salinity8, xmn=min(lat), xmx=max(lat), ymn=min(lon), ymx=max(lon), crs=CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0"))
r_brick9 <- brick(salinity9, xmn=min(lat), xmx=max(lat), ymn=min(lon), ymx=max(lon), crs=CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0"))
r_brick10 <- brick(salinity10, xmn=min(lat), xmx=max(lat), ymn=min(lon), ymx=max(lon), crs=CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0"))
r_brick11 <- brick(salinity11, xmn=min(lat), xmx=max(lat), ymn=min(lon), ymx=max(lon), crs=CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0"))
r_brick12 <- brick(salinity12, xmn=min(lat), xmx=max(lat), ymn=min(lon), ymx=max(lon), crs=CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0"))

r_brick1 <- flip(t(r_brick1), direction='y')
r_brick2 <- flip(t(r_brick2), direction='y')
r_brick3 <- flip(t(r_brick3), direction='y')
r_brick4 <- flip(t(r_brick4), direction='y')
r_brick5 <- flip(t(r_brick5), direction='y')
r_brick6 <- flip(t(r_brick6), direction='y')
r_brick7 <- flip(t(r_brick7), direction='y')
r_brick8 <- flip(t(r_brick8), direction='y')
r_brick9 <- flip(t(r_brick9), direction='y')
r_brick10 <- flip(t(r_brick10), direction='y')
r_brick11 <- flip(t(r_brick11), direction='y')
r_brick12 <- flip(t(r_brick12), direction='y')
```

##get data for each location
```{r}
# traj10year$lon[traj10year$lon > 180] <- traj10year$lon[traj10year$lon > 180] - 360
metadata <- readxl::read_excel("/Users/stephaniematthews/Documents/Chapter4/Organization_Pacific_Basin_Comprehensive_v2.xlsx", sheet = "Locations")
samplelocationcoords <- metadata[,c("Lat(North)", "Lon(West)", "Final Number")]
colnames(samplelocationcoords) <- c("Lat.N.", "Lon.W.", "Location.ID")
  #unique(data.frame(sample_data(zoops18S)[,c("Lat.N.", "Lon.W.", "Location.ID")]))
samplelocationcoords$Lon.E. <- samplelocationcoords$Lon.W. + 360

salinitycalcs <- function(samplelocationcoords) {
  salinityprofiles1 <- NULL
  salinityprofiles2 <- NULL
  salinityprofiles3 <- NULL
  salinityprofiles4 <- NULL
  salinityprofiles5 <- NULL
  salinityprofiles6 <- NULL
  salinityprofiles7 <- NULL
  salinityprofiles8 <- NULL
  salinityprofiles9 <- NULL
  salinityprofiles10 <- NULL
  salinityprofiles11 <- NULL
  salinityprofiles12 <- NULL
  for (samplelocation in 1:nrow(samplelocationcoords)) {
    toolik_lon <- samplelocationcoords[samplelocation,4]
    toolik_lat <- samplelocationcoords[samplelocation,1]
    salinityprofilessingle1 <- raster::extract(r_brick1, SpatialPoints(cbind(toolik_lon,toolik_lat)), method='simple', buffer = 200*1000, fun = mean)#buffer in m for radius around which to extract points
    salinityprofilessingle2 <- raster::extract(r_brick2, SpatialPoints(cbind(toolik_lon,toolik_lat)), method='simple', buffer = 200*1000, fun = mean)
    salinityprofilessingle3 <- raster::extract(r_brick3, SpatialPoints(cbind(toolik_lon,toolik_lat)), method='simple', buffer = 200*1000, fun = mean)
    salinityprofilessingle4 <- raster::extract(r_brick4, SpatialPoints(cbind(toolik_lon,toolik_lat)), method='simple', buffer = 200*1000, fun = mean)
    salinityprofilessingle5 <- raster::extract(r_brick5, SpatialPoints(cbind(toolik_lon,toolik_lat)), method='simple', buffer = 200*1000, fun = mean)
    salinityprofilessingle6 <- raster::extract(r_brick6, SpatialPoints(cbind(toolik_lon,toolik_lat)), method='simple', buffer = 200*1000, fun = mean)
    salinityprofilessingle7 <- raster::extract(r_brick7, SpatialPoints(cbind(toolik_lon,toolik_lat)), method='simple', buffer = 200*1000, fun = mean)
    salinityprofilessingle8 <- raster::extract(r_brick8, SpatialPoints(cbind(toolik_lon,toolik_lat)), method='simple', buffer = 200*1000, fun = mean)
    salinityprofilessingle9 <- raster::extract(r_brick9, SpatialPoints(cbind(toolik_lon,toolik_lat)), method='simple', buffer = 200*1000, fun = mean)
    salinityprofilessingle10 <- raster::extract(r_brick10, SpatialPoints(cbind(toolik_lon,toolik_lat)), method='simple', buffer = 200*1000, fun = mean)
    salinityprofilessingle11 <- raster::extract(r_brick11, SpatialPoints(cbind(toolik_lon,toolik_lat)), method='simple', buffer = 200*1000, fun = mean)
    salinityprofilessingle12 <- raster::extract(r_brick12, SpatialPoints(cbind(toolik_lon,toolik_lat)), method='simple', buffer = 200*1000, fun = mean)
    salinityprofiles1 <- rbind(salinityprofiles1, salinityprofilessingle1)
    salinityprofiles2 <- rbind(salinityprofiles2, salinityprofilessingle2)
    salinityprofiles3 <- rbind(salinityprofiles3, salinityprofilessingle3)
    salinityprofiles4 <- rbind(salinityprofiles4, salinityprofilessingle4)
    salinityprofiles5 <- rbind(salinityprofiles5, salinityprofilessingle5)
    salinityprofiles6 <- rbind(salinityprofiles6, salinityprofilessingle6)
    salinityprofiles7 <- rbind(salinityprofiles7, salinityprofilessingle7)
    salinityprofiles8 <- rbind(salinityprofiles8, salinityprofilessingle8)
    salinityprofiles9 <- rbind(salinityprofiles9, salinityprofilessingle9)
    salinityprofiles10 <- rbind(salinityprofiles10, salinityprofilessingle10)
    salinityprofiles11 <- rbind(salinityprofiles11, salinityprofilessingle11)
    salinityprofiles12 <- rbind(salinityprofiles12, salinityprofilessingle12)
  }
  rownames(salinityprofiles1)  <- paste(samplelocationcoords$Location.ID, "Jan")
  rownames(salinityprofiles2)  <- paste(samplelocationcoords$Location.ID, "Feb")
  rownames(salinityprofiles3)  <- paste(samplelocationcoords$Location.ID, "Mar")
  rownames(salinityprofiles4)  <- paste(samplelocationcoords$Location.ID, "Apr")
  rownames(salinityprofiles5)  <- paste(samplelocationcoords$Location.ID, "May")
  rownames(salinityprofiles6)  <- paste(samplelocationcoords$Location.ID, "Jun")
  rownames(salinityprofiles7)  <- paste(samplelocationcoords$Location.ID, "Jul")
  rownames(salinityprofiles8)  <- paste(samplelocationcoords$Location.ID, "Aug")
  rownames(salinityprofiles9)  <- paste(samplelocationcoords$Location.ID, "Sep")
  rownames(salinityprofiles10) <- paste(samplelocationcoords$Location.ID, "Oct")
  rownames(salinityprofiles11) <- paste(samplelocationcoords$Location.ID, "Nov")
  rownames(salinityprofiles12) <- paste(samplelocationcoords$Location.ID, "Dec")
  
  salinityprofilesall <- rbind(salinityprofiles1, salinityprofiles2) %>% 
    rbind(salinityprofiles3)  %>% 
    rbind(salinityprofiles4)  %>% 
    rbind(salinityprofiles5)  %>% 
    rbind(salinityprofiles6)  %>% 
    rbind(salinityprofiles7)  %>% 
    rbind(salinityprofiles8)  %>% 
    rbind(salinityprofiles9)  %>% 
    rbind(salinityprofiles10) %>% 
    rbind(salinityprofiles11) %>% 
    rbind(salinityprofiles12)

  print(salinityprofilesall)
}

salinityout <- salinitycalcs(samplelocationcoords)
colnames(salinityout) <- levels_m
```

##plot salinity profiles
```{r}
sallong <- data.frame(salinityout) %>% tibble::rownames_to_column(var = "Row") %>% pivot_longer(-Row, names_to = "depth", values_to = "salinity" )
sallong$Location <- stringr::str_split_i((sallong$Row), " ", i = 1)
sallong$Month <- stringr::str_split_i((sallong$Row), " ", i = 2)
sallong$Depth_m <- as.numeric(gsub("X","",as.character(sallong$depth)))
#sallong$Depth_m <- (as.numeric(stringr::str_split_i((sallong$depth), "[.]", i = 2))*111.27)-108.77
sallong$salinity_calc2 <- (sallong$salinity+35)#*1000
sallong$salinity_calc <- (sallong$salinity*1000)+35
  
ggplot(sallong, aes(x=Depth_m, y=salinity_calc, group=Month)) +
  geom_line() + # make this a line plot
  ggtitle("Mean Salinity 200km around each station") +     # Set title
  theme_bw() + # use the black and white theme + 
  coord_flip() +
  facet_wrap(.~Location) + 
  xlim(1000,0)
```



#TEMPERATURE
##get data
```{r}
nc_data <- nc_open('/Users/stephaniematthews/ofesdata/temp00000.nc')
lon <- ncvar_get(nc_data, "lon")
lat <- ncvar_get(nc_data, "lat", verbose = F)
temp1 <- ncvar_get(nc_data, "temp") # store the data in a 3-dimensional array
fillvalue <- ncatt_get(nc_data, "temp", "_FillValue")
levels_m <- ncvar_get(nc_data, "lev")
fillvalue
nc_close(nc_data) 
nc_data <- nc_open('/Users/stephaniematthews/ofesdata/temp00001.nc')
temp2 <- ncvar_get(nc_data, "temp") # store the data in a 3-dimensional array
nc_close(nc_data) 
nc_data <- nc_open('/Users/stephaniematthews/ofesdata/temp00002.nc')
temp3 <- ncvar_get(nc_data, "temp") # store the data in a 3-dimensional array
nc_close(nc_data) 
nc_data <- nc_open('/Users/stephaniematthews/ofesdata/temp00003.nc')
temp4 <- ncvar_get(nc_data, "temp") # store the data in a 3-dimensional array
nc_close(nc_data) 
nc_data <- nc_open('/Users/stephaniematthews/ofesdata/temp00004.nc')
temp5 <- ncvar_get(nc_data, "temp") # store the data in a 3-dimensional array
nc_close(nc_data) 
nc_data <- nc_open('/Users/stephaniematthews/ofesdata/temp00005.nc')
temp6 <- ncvar_get(nc_data, "temp") # store the data in a 3-dimensional array
nc_close(nc_data) 
nc_data <- nc_open('/Users/stephaniematthews/ofesdata/temp00006.nc')
temp7 <- ncvar_get(nc_data, "temp") # store the data in a 3-dimensional array
nc_close(nc_data) 
nc_data <- nc_open('/Users/stephaniematthews/ofesdata/temp00007.nc')
temp8 <- ncvar_get(nc_data, "temp") # store the data in a 3-dimensional array
nc_close(nc_data) 
nc_data <- nc_open('/Users/stephaniematthews/ofesdata/temp00008.nc')
temp9 <- ncvar_get(nc_data, "temp") # store the data in a 3-dimensional array
nc_close(nc_data) 
nc_data <- nc_open('/Users/stephaniematthews/ofesdata/temp00009.nc')
temp10 <- ncvar_get(nc_data, "temp") # store the data in a 3-dimensional array
nc_close(nc_data) 
nc_data <- nc_open('/Users/stephaniematthews/ofesdata/temp00010.nc')
temp11 <- ncvar_get(nc_data, "temp") # store the data in a 3-dimensional array
nc_close(nc_data) 
nc_data <- nc_open('/Users/stephaniematthews/ofesdata/temp00011.nc')
temp12 <- ncvar_get(nc_data, "temp") # store the data in a 3-dimensional array
nc_close(nc_data) 
```

##reassign NA value
```{r}
temp1[temp1 == fillvalue$value] <- NA
temp2[temp2 == fillvalue$value] <- NA
temp3[temp3 == fillvalue$value] <- NA
temp4[temp4 == fillvalue$value] <- NA
temp5[temp5 == fillvalue$value] <- NA
temp6[temp6 == fillvalue$value] <- NA
temp7[temp7 == fillvalue$value] <- NA
temp8[temp8 == fillvalue$value] <- NA
temp9[temp9 == fillvalue$value] <- NA
temp10[temp10 == fillvalue$value] <- NA
temp11[temp11 == fillvalue$value] <- NA
temp12[temp12 == fillvalue$value] <- NA
```

##format as brick
```{r}
r_brick1 <- brick(temp1, xmn=min(lat), xmx=max(lat), ymn=min(lon), ymx=max(lon), crs=CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0"))
r_brick2 <- brick(temp2, xmn=min(lat), xmx=max(lat), ymn=min(lon), ymx=max(lon), crs=CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0"))
r_brick3 <- brick(temp3, xmn=min(lat), xmx=max(lat), ymn=min(lon), ymx=max(lon), crs=CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0"))
r_brick4 <- brick(temp4, xmn=min(lat), xmx=max(lat), ymn=min(lon), ymx=max(lon), crs=CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0"))
r_brick5 <- brick(temp5, xmn=min(lat), xmx=max(lat), ymn=min(lon), ymx=max(lon), crs=CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0"))
r_brick6 <- brick(temp6, xmn=min(lat), xmx=max(lat), ymn=min(lon), ymx=max(lon), crs=CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0"))
r_brick7 <- brick(temp7, xmn=min(lat), xmx=max(lat), ymn=min(lon), ymx=max(lon), crs=CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0"))
r_brick8 <- brick(temp8, xmn=min(lat), xmx=max(lat), ymn=min(lon), ymx=max(lon), crs=CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0"))
r_brick9 <- brick(temp9, xmn=min(lat), xmx=max(lat), ymn=min(lon), ymx=max(lon), crs=CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0"))
r_brick10 <- brick(temp10, xmn=min(lat), xmx=max(lat), ymn=min(lon), ymx=max(lon), crs=CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0"))
r_brick11 <- brick(temp11, xmn=min(lat), xmx=max(lat), ymn=min(lon), ymx=max(lon), crs=CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0"))
r_brick12 <- brick(temp12, xmn=min(lat), xmx=max(lat), ymn=min(lon), ymx=max(lon), crs=CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0"))

r_brick1 <- flip(t(r_brick1), direction='y')
r_brick2 <- flip(t(r_brick2), direction='y')
r_brick3 <- flip(t(r_brick3), direction='y')
r_brick4 <- flip(t(r_brick4), direction='y')
r_brick5 <- flip(t(r_brick5), direction='y')
r_brick6 <- flip(t(r_brick6), direction='y')
r_brick7 <- flip(t(r_brick7), direction='y')
r_brick8 <- flip(t(r_brick8), direction='y')
r_brick9 <- flip(t(r_brick9), direction='y')
r_brick10 <- flip(t(r_brick10), direction='y')
r_brick11 <- flip(t(r_brick11), direction='y')
r_brick12 <- flip(t(r_brick12), direction='y')
```

##get data for each location
```{r}
# traj10year$lon[traj10year$lon > 180] <- traj10year$lon[traj10year$lon > 180] - 360
metadata <- readxl::read_excel("/Users/stephaniematthews/Documents/Chapter4/Organization_Pacific_Basin_Comprehensive_v2.xlsx", sheet = "Locations")
samplelocationcoords <- metadata[,c("Lat(North)", "Lon(West)", "Final Number")]
colnames(samplelocationcoords) <- c("Lat.N.", "Lon.W.", "Location.ID")
  #unique(data.frame(sample_data(zoops18S)[,c("Lat.N.", "Lon.W.", "Location.ID")]))
samplelocationcoords$Lon.E. <- samplelocationcoords$Lon.W. + 360

tempcalcs <- function(samplelocationcoords) {
  tempprofiles1 <- NULL
  tempprofiles2 <- NULL
  tempprofiles3 <- NULL
  tempprofiles4 <- NULL
  tempprofiles5 <- NULL
  tempprofiles6 <- NULL
  tempprofiles7 <- NULL
  tempprofiles8 <- NULL
  tempprofiles9 <- NULL
  tempprofiles10 <- NULL
  tempprofiles11 <- NULL
  tempprofiles12 <- NULL
  for (samplelocation in 1:nrow(samplelocationcoords)) {
    toolik_lon <- samplelocationcoords[samplelocation,4]
    toolik_lat <- samplelocationcoords[samplelocation,1]
    tempprofilessingle1 <- raster::extract(r_brick1, SpatialPoints(cbind(toolik_lon,toolik_lat)), method='simple', buffer = 200*1000, fun = mean)#buffer in m for radius around which to extract points
    tempprofilessingle2 <- raster::extract(r_brick2, SpatialPoints(cbind(toolik_lon,toolik_lat)), method='simple', buffer = 200*1000, fun = mean)
    tempprofilessingle3 <- raster::extract(r_brick3, SpatialPoints(cbind(toolik_lon,toolik_lat)), method='simple', buffer = 200*1000, fun = mean)
    tempprofilessingle4 <- raster::extract(r_brick4, SpatialPoints(cbind(toolik_lon,toolik_lat)), method='simple', buffer = 200*1000, fun = mean)
    tempprofilessingle5 <- raster::extract(r_brick5, SpatialPoints(cbind(toolik_lon,toolik_lat)), method='simple', buffer = 200*1000, fun = mean)
    tempprofilessingle6 <- raster::extract(r_brick6, SpatialPoints(cbind(toolik_lon,toolik_lat)), method='simple', buffer = 200*1000, fun = mean)
    tempprofilessingle7 <- raster::extract(r_brick7, SpatialPoints(cbind(toolik_lon,toolik_lat)), method='simple', buffer = 200*1000, fun = mean)
    tempprofilessingle8 <- raster::extract(r_brick8, SpatialPoints(cbind(toolik_lon,toolik_lat)), method='simple', buffer = 200*1000, fun = mean)
    tempprofilessingle9 <- raster::extract(r_brick9, SpatialPoints(cbind(toolik_lon,toolik_lat)), method='simple', buffer = 200*1000, fun = mean)
    tempprofilessingle10 <- raster::extract(r_brick10, SpatialPoints(cbind(toolik_lon,toolik_lat)), method='simple', buffer = 200*1000, fun = mean)
    tempprofilessingle11 <- raster::extract(r_brick11, SpatialPoints(cbind(toolik_lon,toolik_lat)), method='simple', buffer = 200*1000, fun = mean)
    tempprofilessingle12 <- raster::extract(r_brick12, SpatialPoints(cbind(toolik_lon,toolik_lat)), method='simple', buffer = 200*1000, fun = mean)
    tempprofiles1 <- rbind(tempprofiles1, tempprofilessingle1)
    tempprofiles2 <- rbind(tempprofiles2, tempprofilessingle2)
    tempprofiles3 <- rbind(tempprofiles3, tempprofilessingle3)
    tempprofiles4 <- rbind(tempprofiles4, tempprofilessingle4)
    tempprofiles5 <- rbind(tempprofiles5, tempprofilessingle5)
    tempprofiles6 <- rbind(tempprofiles6, tempprofilessingle6)
    tempprofiles7 <- rbind(tempprofiles7, tempprofilessingle7)
    tempprofiles8 <- rbind(tempprofiles8, tempprofilessingle8)
    tempprofiles9 <- rbind(tempprofiles9, tempprofilessingle9)
    tempprofiles10 <- rbind(tempprofiles10, tempprofilessingle10)
    tempprofiles11 <- rbind(tempprofiles11, tempprofilessingle11)
    tempprofiles12 <- rbind(tempprofiles12, tempprofilessingle12)
  }
  rownames(tempprofiles1)  <- paste(samplelocationcoords$Location.ID, "Jan")
  rownames(tempprofiles2)  <- paste(samplelocationcoords$Location.ID, "Feb")
  rownames(tempprofiles3)  <- paste(samplelocationcoords$Location.ID, "Mar")
  rownames(tempprofiles4)  <- paste(samplelocationcoords$Location.ID, "Apr")
  rownames(tempprofiles5)  <- paste(samplelocationcoords$Location.ID, "May")
  rownames(tempprofiles6)  <- paste(samplelocationcoords$Location.ID, "Jun")
  rownames(tempprofiles7)  <- paste(samplelocationcoords$Location.ID, "Jul")
  rownames(tempprofiles8)  <- paste(samplelocationcoords$Location.ID, "Aug")
  rownames(tempprofiles9)  <- paste(samplelocationcoords$Location.ID, "Sep")
  rownames(tempprofiles10) <- paste(samplelocationcoords$Location.ID, "Oct")
  rownames(tempprofiles11) <- paste(samplelocationcoords$Location.ID, "Nov")
  rownames(tempprofiles12) <- paste(samplelocationcoords$Location.ID, "Dec")
  
  tempprofilesall <- rbind(tempprofiles1, tempprofiles2) %>% 
    rbind(tempprofiles3)  %>% 
    rbind(tempprofiles4)  %>% 
    rbind(tempprofiles5)  %>% 
    rbind(tempprofiles6)  %>% 
    rbind(tempprofiles7)  %>% 
    rbind(tempprofiles8)  %>% 
    rbind(tempprofiles9)  %>% 
    rbind(tempprofiles10) %>% 
    rbind(tempprofiles11) %>% 
    rbind(tempprofiles12)

  print(tempprofilesall)
}

tempout <- tempcalcs(samplelocationcoords)
colnames(tempout) <- levels_m

```

##plot temperature profiles
```{r}
tlong <- data.frame(tempout) %>% tibble::rownames_to_column(var = "Row") %>% pivot_longer(-Row, names_to = "depth", values_to = "temp" )
tlong$Location <- stringr::str_split_i((tlong$Row), " ", i = 1)
tlong$Month <- stringr::str_split_i((tlong$Row), " ", i = 2)
tlong$Depth_m <- as.numeric(gsub("X","",as.character(sallong$depth)))
#tlong$Depth_m <- (as.numeric(stringr::str_split_i((tlong$depth), "[.]", i = 2))*111.27)-108.77
#sallong$temp_calc <- (sallong$temp+35)#*1000
  
ggplot(tlong, aes(x=Depth_m, y=temp, group=Month)) +
  geom_line() + # make this a line plot
  ggtitle("Mean Temp 200km around each station") +     # Set title
  theme_bw() + # use the black and white theme + 
  coord_flip() +
  facet_wrap(.~Location) + 
  xlim(1000,0)
```



#calculate means 
```{r}
tlong$net <- NA
tlong$net[tlong$Depth_m %in% c(2.50000,    7.56156,   12.80628,   18.35124,   24.30616,   30.77062,   37.83170,   45.56174,   54.01666,   63.23449,   73.23449,   84.01666,   95.56174,  107.83170,  120.77062,  134.30615,  148.35124,  162.80627,  177.56155,  192.50000)] <- "0-200m"
tlong$net[tlong$Depth_m %in% c(207.50000,  223.17186,  240.85361,  261.86606,  287.50168,  319.01346,  357.60428,  404.41656,  460.52264)] <- "200-500m"
tlong$net[tlong$Depth_m %in% c(526.91559,  604.50061,  694.08759,  796.38379,  911.98822, 1041.38635)] <- "500-1000m"
tlong <- tlong[tlong$Depth_m < 1050,]

sallong$net <- NA
sallong$net[sallong$Depth_m %in% c(2.50000,    7.56156,   12.80628,   18.35124,   24.30616,   30.77062,   37.83170,   45.56174,   54.01666,   63.23449,   73.23449,   84.01666,   95.56174,  107.83170,  120.77062,  134.30615,  148.35124,  162.80627,  177.56155,  192.50000)] <- "0-200m"
sallong$net[sallong$Depth_m %in% c(207.50000,  223.17186,  240.85361,  261.86606,  287.50168,  319.01346,  357.60428,  404.41656,  460.52264)] <- "200-500m"
sallong$net[sallong$Depth_m %in% c(558.85,670.12, 781.39, 892.66, 1003.93)] <- "500-1000m"
sallong$net[sallong$Depth_m %in% c(526.91559,  604.50061,  694.08759,  796.38379,  911.98822, 1041.38635)] <- "500-1000m"
sallong <- sallong[sallong$Depth_m < 1050,]


tmeans <- tlong %>% dplyr::summarize(.by = c("net", "Location", "Month"), mean = mean(temp))
smeans <- sallong %>% dplyr::summarize(.by = c("net", "Location", "Month"), mean = mean(salinity_calc))

tmeanswide <- tmeans %>% pivot_wider(names_from = c(Location, net), values_from = mean, names_glue = "{Location} {net}") %>% t()
smeanswide <- smeans %>% pivot_wider(names_from = c(Location, net), values_from = mean, names_glue = "{Location} {net}") %>% t()

colnames(tmeanswide) <- paste("Temp, ", tmeanswide[1,])
colnames(smeanswide) <- paste("Sal, ", smeanswide[1,])

tmeanswide <- data.frame(tmeanswide[-1,])
smeanswide <- data.frame(smeanswide[-1,])
```


#COI
#merge with sample metadata
```{r}
insitu <- data.frame(sample_data(zoopsCOI))
insitu <- insitu[,c(1:41)]
insitu$uniqueidentifier <- sample_names(zoopsCOI)
insitu$formerging <- paste(insitu$StationID, insitu$depth_bin)

tmeanswide$formerging <- rownames(tmeanswide)
smeanswide$formerging <- rownames(smeanswide)

insituandmodel <- merge(insitu, tmeanswide, by = "formerging", all.x = T, all.y = F)
insituandmodel <- merge(insituandmodel, smeanswide, by = "formerging", all.x = T, all.y = F)

```

#plot in situ vs model
```{r}

sal3depths <- insituandmodel %>% dplyr::select(starts_with("Sal..."), ends_with("Sal"), depth_bin)

temp3depths <- insituandmodel %>% dplyr::select(starts_with("Temp..."), ends_with("Temp"), depth_bin)

#tlong <- data.frame(tempout) %>% tibble::rownames_to_column(var = "Row") %>% pivot_longer(-Row, names_to = "depth", values_to = "temp" ) %dplyr::select(!depth_bin) %>%
salcomparison <- sal3depths %>%  pivot_longer(-c(MaxSal, MeanSal, MinSal, depth_bin), names_to = "Month", values_to = "ModelSalinity") %>% dplyr::mutate_at(c('ModelSalinity'), as.numeric) %>% ggplot(aes(x = MeanSal, y = ModelSalinity)) + geom_point() + facet_wrap(.~depth_bin, scales = "free") + geom_abline(slope=1, intercept=0) + xlab("Measured Salinity (mean)") + ylab("Modeled Salinity\n(monthly mean)") 

jpeg(file = "~/Documents/Chapter4/BasinScale_Analysis_v2_2024/figures/Salinity_climatology_vs_insitu.jpeg", height =2, width = 6.5, unit = "in", quality = 100, res = 300)
salcomparison
dev.off()

tempcomparison <- temp3depths %>% pivot_longer(-c(MaxTemp, MeanTemp, MinTemp, depth_bin), names_to = "Month", values_to = "ModelTemperature") %>% dplyr::mutate_at(c('ModelTemperature'), as.numeric) %>% ggplot(aes(x = MeanTemp, y = ModelTemperature)) + geom_point() + facet_wrap(.~depth_bin, scales = "free") + geom_abline(slope=1, intercept=0)+ xlab("Measured Temperature (mean)") + ylab("Modeled Temperature\n(monthly mean)")

jpeg(file = "~/Documents/Chapter4/BasinScale_Analysis_v2_2024/figures/Temp_climatology_vs_insitu.jpeg", height =2, width = 6.5, unit = "in", quality = 100, res = 300)
tempcomparison
dev.off()

```

#calculate annual amplitude and means
```{r}
insituandmodel <- insituandmodel %>% mutate_at(c("Temp...Jan", "Temp...Feb", "Temp...Mar", "Temp...Apr", "Temp...May", "Temp...Jun", "Temp...Jul", "Temp...Aug", "Temp...Sep", "Temp...Oct", "Temp...Nov", "Temp...Dec", "Sal...Jan", "Sal...Feb", "Sal...Mar", "Sal...Apr", "Sal...May", "Sal...Jun", "Sal...Jul", "Sal...Aug", "Sal...Sep", "Sal...Oct", "Sal...Nov", "Sal...Dec"), as.numeric)

insituandmodel$meanModelTemp <- rowMeans(insituandmodel[,c("Temp...Jan", "Temp...Feb", "Temp...Mar", "Temp...Apr", "Temp...May", "Temp...Jun", "Temp...Jul", "Temp...Aug", "Temp...Sep", "Temp...Oct", "Temp...Nov", "Temp...Dec")])
insituandmodel$meanModelSal <- rowMeans(insituandmodel[,c("Sal...Jan", "Sal...Feb", "Sal...Mar", "Sal...Apr", "Sal...May", "Sal...Jun", "Sal...Jul", "Sal...Aug", "Sal...Sep", "Sal...Oct", "Sal...Nov", "Sal...Dec")])
insituandmodel <- insituandmodel %>% 
  rowwise() %>% 
  mutate(maxModelTemp = as.numeric(max(c_across(starts_with("Temp...")))), minModelTemp = as.numeric(min(c_across(starts_with("Temp...")))),
         maxModelSal = as.numeric(max(c_across(starts_with("Sal...")))), minModelSal = as.numeric(min(c_across(starts_with("Sal...")))))
insituandmodel$ampModelTemp <- insituandmodel$maxModelTemp - insituandmodel$minModelTemp
insituandmodel$ampModelSal <- insituandmodel$maxModelSal - insituandmodel$minModelSal
insituandmodel <- data.frame(insituandmodel)

insituandmodel <- insituandmodel[!is.na(insituandmodel$depth_bin),]
rownames(insituandmodel) <- insituandmodel$uniqueidentifier
sample_data(zoopsCOI) <- sample_data(insituandmodel)

```


#18S
#merge with sample metadata
```{r}
insitu <- data.frame(sample_data(zoops18S))
insitu <- insitu[,c(1:41)]
insitu$uniqueidentifier <- sample_names(zoops18S)
insitu$formerging <- paste(insitu$StationID, insitu$depth_bin)

tmeanswide$formerging <- rownames(tmeanswide)
smeanswide$formerging <- rownames(smeanswide)

insituandmodel <- merge(insitu, tmeanswide, by = "formerging", all.x = T, all.y = F)
insituandmodel <- merge(insituandmodel, smeanswide, by = "formerging", all.x = T, all.y = F)

```

#plot in situ vs model
```{r}

sal3depths <- insituandmodel %>% dplyr::select(starts_with("Sal..."), ends_with("Sal"), depth_bin)
salepi <- sal3depths[sal3depths$depth_bin == "0-200m",]
salupme <- sal3depths[sal3depths$depth_bin == "200-500m",]
sallowme <- sal3depths[sal3depths$depth_bin == "500-1000m",]

temp3depths <- insituandmodel %>% dplyr::select(starts_with("Temp..."), ends_with("Temp"), depth_bin)
tempepi <- temp3depths[temp3depths$depth_bin == "0-200m",]
tempupme <- temp3depths[temp3depths$depth_bin == "200-500m",]
templowme <- temp3depths[temp3depths$depth_bin == "500-1000m",]

#tlong <- data.frame(tempout) %>% tibble::rownames_to_column(var = "Row") %>% pivot_longer(-Row, names_to = "depth", values_to = "temp" ) %dplyr::select(!depth_bin) %>%
sal3depths %>%  pivot_longer(-c(MaxSal, MeanSal, MinSal, depth_bin), names_to = "Month", values_to = "ModelSalinity") %>% dplyr::mutate_at(c('ModelSalinity'), as.numeric) %>% ggplot(aes(x = MeanSal, y = ModelSalinity)) + geom_point() + facet_wrap(.~depth_bin, scales = "free")+ geom_abline(slope=1, intercept=0)

temp3depths %>%  pivot_longer(-c(MaxTemp, MeanTemp, MinTemp, depth_bin), names_to = "Month", values_to = "ModelTemperature") %>% dplyr::mutate_at(c('ModelTemperature'), as.numeric) %>% ggplot(aes(x = MeanTemp, y = ModelTemperature)) + geom_point() + facet_wrap(.~depth_bin, scales = "free")+ geom_abline(slope=1, intercept=0)

```

#calculate annual amplitude and means
```{r}
insituandmodel <- insituandmodel %>% mutate_at(c("Temp...Jan", "Temp...Feb", "Temp...Mar", "Temp...Apr", "Temp...May", "Temp...Jun", "Temp...Jul", "Temp...Aug", "Temp...Sep", "Temp...Oct", "Temp...Nov", "Temp...Dec", "Sal...Jan", "Sal...Feb", "Sal...Mar", "Sal...Apr", "Sal...May", "Sal...Jun", "Sal...Jul", "Sal...Aug", "Sal...Sep", "Sal...Oct", "Sal...Nov", "Sal...Dec"), as.numeric)

insituandmodel$meanModelTemp <- rowMeans(insituandmodel[,c("Temp...Jan", "Temp...Feb", "Temp...Mar", "Temp...Apr", "Temp...May", "Temp...Jun", "Temp...Jul", "Temp...Aug", "Temp...Sep", "Temp...Oct", "Temp...Nov", "Temp...Dec")])
insituandmodel$meanModelSal <- rowMeans(insituandmodel[,c("Sal...Jan", "Sal...Feb", "Sal...Mar", "Sal...Apr", "Sal...May", "Sal...Jun", "Sal...Jul", "Sal...Aug", "Sal...Sep", "Sal...Oct", "Sal...Nov", "Sal...Dec")])
insituandmodel <- insituandmodel %>% 
  rowwise() %>% 
  mutate(maxModelTemp = as.numeric(max(c_across(starts_with("Temp...")))), minModelTemp = as.numeric(min(c_across(starts_with("Temp...")))),
         maxModelSal = as.numeric(max(c_across(starts_with("Sal...")))), minModelSal = as.numeric(min(c_across(starts_with("Sal...")))))
insituandmodel$ampModelTemp <- insituandmodel$maxModelTemp - insituandmodel$minModelTemp
insituandmodel$ampModelSal <- insituandmodel$maxModelSal - insituandmodel$minModelSal
insituandmodel <- data.frame(insituandmodel)

insituandmodel <- insituandmodel[!is.na(insituandmodel$depth_bin),]
rownames(insituandmodel) <- insituandmodel$uniqueidentifier
sample_data(zoops18S) <- sample_data(insituandmodel)
```

```{r}
#save(zoops18S, zoopsCOI, file = "~/Documents/Chapter4/BasinScale_AllSamples_v1/CleanZooplanktonPhyloseqs.rdat")

```


```{r}
```












