---
title: "Analyze_07_Mantel"
output: html_document
date: "2024-11-19"
---


```{r setup, include=FALSE}
library(microshades)
library(phyloseq)
library(vegan)
library(ggplot2)
library(dplyr)
library(tidyr)
library(metagMisc)
#library(viridis)
library(RColorBrewer)
library(ggExtra)
library(beepr)
library(readxl)
library(ggforce)
library(ggrepel)
library(ggpubr)
library(cowplot)
library(ggtext)
library(sp)
library(vegan)
library(grid)
library(ggsci)
set.seed(00110011)
theme_set(theme_bw())
theme_update(plot.title = element_text(hjust = 0.5))
depthcolors <- c("#4DAF4A", "#E41A1C", "#984EA3")
names(depthcolors) <- c("0-200m", "200-500m", "500-1000m")
markercolors <- c("#D81B60","#1E88E5")
names(markercolors) <- c("18S", "COI")

```

```{r}
load("~/Documents/Chapter4/BasinScale_Analysis_v2_2024/data_objects/CleanZooplanktonPhyloseqs.rdat")
```


# environment similarity:community similarity
- Does epipelagic/mesopelagic community similarity covary with environmental similarity? 
- Can I calculate distance among samples and test for correlation with environmental differences? 
- Is environmental similarity the same as biogeochemical provinces? Consider whether to include biogeochemical provinces

-> we'll use bray-curtis similarity, since there's variability in sampling type and depth bins etc 
##get distance matrixes, 18S
```{r community distance}
#subset to depths
epi18 <- subset_samples(zoops18S, depthbin == "0-200m")
upmeso18 <- subset_samples(zoops18S, depthbin == "200-500m")
lowmeso18 <- subset_samples(zoops18S, depthbin == "500-1000m")

#bray curtis
bc.epi18 <- phyloseq::distance(epi18, method = "bray")
bc.upmeso18 <- phyloseq::distance(upmeso18, method = "bray")
bc.lowmeso18 <- phyloseq::distance(lowmeso18, method = "bray")

#model temp
epi18.modtemp <- as(sample_data(epi18), "data.frame") %>% dplyr::select(ends_with("meanModelTemp")) %>% dist(method = "maximum")
upmeso18.modtemp <- as(sample_data(upmeso18), "data.frame") %>% dplyr::select("meanModelTemp") %>% dist(method = "maximum")
lowmeso18.modtemp <- as(sample_data(lowmeso18), "data.frame") %>% dplyr::select("meanModelTemp") %>% dist(method = "maximum")

#model salinity
epi18.modsal <- as(sample_data(epi18), "data.frame") %>% dplyr::select(ends_with("meanModelSal")) %>% dist(method = "maximum")
upmeso18.modsal <- as(sample_data(upmeso18), "data.frame") %>% dplyr::select("meanModelSal") %>% dist(method = "maximum")
lowmeso18.modsal <- as(sample_data(lowmeso18), "data.frame") %>% dplyr::select("meanModelSal") %>% dist(method = "maximum")

#measured temp
epi18.temp <- as(sample_data(epi18), "data.frame") %>% dplyr::select(ends_with("MeanTemp")) %>% dist(method = "maximum")
upmeso18.temp <- as(sample_data(upmeso18), "data.frame") %>% dplyr::select("MeanTemp") %>% dist(method = "maximum")
lowmeso18.temp <- as(sample_data(lowmeso18), "data.frame") %>% dplyr::select("MeanTemp") %>% dist(method = "maximum")

#measured salinity
epi18.sal <- as(sample_data(epi18), "data.frame") %>% dplyr::select(ends_with("MeanSal")) %>% dist(method = "maximum")
upmeso18.sal <- as(sample_data(upmeso18), "data.frame") %>% dplyr::select("MeanSal") %>% dist(method = "maximum")
lowmeso18.sal <- as(sample_data(lowmeso18), "data.frame") %>% dplyr::select("MeanSal") %>% dist(method = "maximum")

#measured oxygen
epi18.oxy <- as(sample_data(epi18), "data.frame") %>% dplyr::select(ends_with("MeanOxy")) %>% dist(method = "maximum")
upmeso18.oxy <- as(sample_data(upmeso18), "data.frame") %>% dplyr::select("MeanOxy") %>% dist(method = "maximum")
lowmeso18.oxy <- as(sample_data(lowmeso18), "data.frame") %>% dplyr::select("MeanOxy") %>% dist(method = "maximum")

#NPP
epi18.npp <- as(sample_data(epi18), "data.frame") %>% dplyr::select(ends_with("npp_meanyearlytotal")) %>% dist(method = "maximum")
upmeso18.npp <- as(sample_data(upmeso18), "data.frame") %>% dplyr::select("npp_meanyearlytotal") %>% dist(method = "maximum")
lowmeso18.npp <- as(sample_data(lowmeso18), "data.frame") %>% dplyr::select("npp_meanyearlytotal") %>% dist(method = "maximum")

#EF
epi18.ef <- as(sample_data(epi18), "data.frame") %>% dplyr::select(ends_with("ef_meanyearlytotal")) %>% dist(method = "maximum")
upmeso18.ef <- as(sample_data(upmeso18), "data.frame") %>% dplyr::select("ef_meanyearlytotal") %>% dist(method = "maximum")
lowmeso18.ef <- as(sample_data(lowmeso18), "data.frame") %>% dplyr::select("ef_meanyearlytotal") %>% dist(method = "maximum")



out <- mantel(epi18.modtemp, bc.epi18, method = "spearman")
mantel(bc.upmeso18, upmeso18.temp, method = "spearman")
mantel(bc.lowmeso18, lowmeso18.temp, method = "spearman")


data.frame(cbind(epi18.temp, bc.epi18)) %>% 
  ggplot(aes(x=epi18.temp, y=bc.epi18)) +
  geom_point(alpha = 0.4, size = .3) + 
#  geom_point()  +theme_bw()+ theme_classic() + 
#  stat_smooth(method = lm, se = T, color = "black") + 
#  geom_point() + theme(text = element_text(size = 16)) 
#  facet_grid(~name_fac, scales = "free_x", labeller = label_parsed) +
  #xlab("Environmental difference") + 
  xlab(expression(paste(Delta, " Environment"))) + 
  ylab("Bray-Curtis Dissimilarity (BCD)") + 
  ggtitle("COI")+ 
   stat_cor(method = "spearman", label.x.npc = .35, label.y.npc = 0.75,
            cor.coef.name = "rho",
            alternative = "greater", geom = "label", 
          #  aes(label = paste(after_stat(r), cut(after_stat(p), 
          #                                    breaks = c(-Inf, 0.001, 0.01, 0.05, Inf),
          #                                    labels = c("'***'", "'**'", "'*'", "'ns'")), sep = "~','~") ),
            show.legend = F) + 
    geom_smooth(method = "loess",  show.legend=F) + 
#  scale_color_manual(values = depthcolors, name = "Depth range")+
 #    guides(color = guide_legend(override.aes = list(shape = c(16, 16, 16), alpha = 1, size = 2) ) )  + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```

```{r}
dist_bray_COI <- phyloseq::distance(zoopsCOI, method = "bray")

```

```{r}
###time matrix
###this makes a matrix with differences between each sample in time
df_md_18 <- as(sample_data(zoops18S), "data.frame")
temp.dist.18 <- df_md_18 %>% dplyr::select(ends_with("meanModelTemp")) %>% dist(method = "maximum")
temp.matrix.18 <- as.matrix(temp.dist.18)
```


```{r}
mantel(dist_bray_18, temp.dist.18, method = "spearman")


```


```{r}


dataP = data.frame(cbind(dist_bray_18, temp.dist.18))


library(ggplot2)
p <- ggplot(dataP, aes(BrayForPlotVir, BrayForPlotBac)) +
  geom_point()  +theme_bw()+ theme_classic() + 
  stat_smooth(method = lm, se = T, color = "black") + 
  geom_point() + theme(text = element_text(size = 16)) + 
  xlab("\nDistance between pairs in viral ASVs (Bray)") +
  ylab("Distance between pairs in bacterial ASVs (Bray)\n")


p


```

